Single Bivalent Level Analysis

### 2020-01-24 doc/Bivalent_Analysis.Rmd
first run of results

                    "F1"='grey', "other" = 'grey'
                    )
```

@@

@@

### Male polymorphism Predcition

@@

```{r male.poly.test.MM, echo=FALSE, eval=FALSE}
#t.test comparisons of PWD_SC vs (KAZ, SKIVE, )
#MSM_SC vs MOLF_SC

@@

rand_SC.Length_MM.male.poly <- exactRLRT(lmer_MM_SC_length_male.poly)#this won't work either...hmm

@@

```{r logreg, echo=FALSE}

#divide musc males into groups
mouse.avs_4MM_male$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM_male$strain), 1, 
                                    ifelse(grepl("MSM", mouse.avs_4MM_male$strain), 1, 0))

log.reg_SC.length_Musc <- glm(Rec.group ~ mean_SC, 
                            data=mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",], family=binomial(link="logit"))

summary(log.reg_SC.length_Musc)

#mean SC slightly sig-- these are all mean SC--should I try distinguishing by chrm classes?
```

```{r long.sc_log.res_meanSC}

@@

#divide musc males into groups
long.SC.mouse.avs_4MM_male$Rec.group <- ifelse(grepl("PWD", long.SC.mouse.avs_4MM_male$strain), 1, 
                                    ifelse(grepl("MSM", long.SC.mouse.avs_4MM_male$strain), 1, 0))

@@

@@

```{r blah.again, echo=FALSE, eval=FALSE}
#long SC Long_biv_mouse.avs_4MM
#real.long.bivData
#Only use 2CO bivalents
real.long.bivData.22 <- real.long.bivData[real.long.bivData$hand.foci.count == 2,]
real.long.bivData.22 <- real.long.bivData.22[(!is.na(real.long.bivData.22$hand.foci.count)),]

@@

@@

```{r IFD.t.test, echo=FALSE, eval=FALSE}

#pool the whole groups

ABS.t.test <- t.test( mouse.avs_4MM_M.male$mean_IFD.2CO_ABS[mouse.avs_4MM_M.male$Rec.group == 1],
  mouse.avs_4MM_M.male$mean_IFD.2CO_ABS[mouse.avs_4MM_M.male$Rec.group == 0])

PER.t.test <- t.test( mouse.avs_4MM_M.male$mean_IFD.2CO_PER[mouse.avs_4MM_M.male$Rec.group == 1],
  mouse.avs_4MM_M.male$mean_IFD.2CO_PER[mouse.avs_4MM_M.male$Rec.group == 0])

#compare strain by strain
# p = `r ABS.t.test$p.value`  and p = `r PER.t.test$p.value`.
```

```{r compare.IFD,echo=FALSE}
#make t.tests for pooled high and low
mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
                                    ifelse(grepl("MSM", mouse.avs_4MM$strain), 1, 0))

Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
                                    ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))

Curated_BivData <- add_IFD(Curated_BivData)

Rec.group_IFD.ABS_test <- t.test(Curated_BivData$IFD1_ABS[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc") & (Curated_BivData$Rec.group ==0 )],
     Curated_BivData$IFD1_ABS[ ( Curated_BivData$Rec.group ==1 )]
)#p-value = 7.955e-05

Rec.group_IFD.ABS_test$p.value

Rec.group_IFD.PER_test <- t.test( Curated_BivData$IFD1_PER[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc") & (Curated_BivData$Rec.group ==0 )],
     Curated_BivData$IFD1_PER[ ( Curated_BivData$Rec.group ==1 )]
)#p-value = 0.04231

@@

biv.IFDs <- ggplot(Curated_BivData[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc"),], aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS")+
  scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,100))

oo.ifd <- ggplot(mouse.avs_4MM[ ( mouse.avs_4MM$sex == "male" ) & (mouse.avs_4MM$subsp == "Musc"),], aes(y = mean_IFD.2CO_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS")+
  scale_color_manual(values=colors_of_strains)+ facet_wrap(~strain, scales = "free")+ylim(c(0,1))+theme(legend.position="none")

@@

@@

@@

> is the male polymorphism

@@

```{r cent.dist, echo=FALSE}
cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))

@@

>>>>

                    "F1"='grey', "other" = 'grey'
                    )

#add Rec group
Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
                                    ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))

@@

```{r XX.adj, echo=FALSE}

#load and examine the the whole cell hand measures here

@@

@@

```{r log.reg_SC.length, echo=FALSE}

#divide musc males into groups
mouse.avs_4MM_male$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM_male$strain), 1, 
                                    ifelse(grepl("MSM", mouse.avs_4MM_male$strain), 1, 0))
#

log.reg_SC.length_Musc <- glm(Rec.group ~ mean_SC, 
                            data=mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",], family=binomial(link="logit"))
mouse.av_log.reg_SC.length <- summary(log.reg_SC.length_Musc)

# bivalent level log regression
log.reg_SC.length_Musc_BIV <- glm(Rec.group ~ chromosomeLength, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male"),], family=binomial(link="logit"))

bivalent_log.reg_SC.length <- summary(log.reg_SC.length_Musc_BIV)

The mean SC logistic regresion model for mouse averages

`r mouse.av_log.reg_SC.length$coefficients`

and for the single bivalent levels

`r bivalent_log.reg_SC.length$coefficients`

#### t.test for sc length

```{r t.test, echo=FALSE}

T.test_Poly_SC.Len_mouse.av <- t.test(mouse.avs_4MM_male$mean_SC[mouse.avs_4MM_male$Rec.group == 1],

  mouse.avs_4MM_male$mean_SC[(mouse.avs_4MM_male$subsp == "Musc")&(mouse.avs_4MM_male$sex == "male")&(mouse.avs_4MM_male$Rec.group == 0)]
)#p-value = 0.001567

T.test_Poly_SC.Len_single.biv <- t.test(Curated_BivData$chromosomeLength[Curated_BivData$Rec.group == 1],

  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)]
)#p-value < 2.2e-16

The mouse average and single bivalents are significantly longer in the high rec group p = `r T.test_Poly_SC.Len_mouse.av$p.value` and p=`r T.test_Poly_SC.Len_single.biv$p.value`.

```{r male.poly.test.MM, echo=FALSE, eval=FALSE}
#I don't think the Mixed models fit well for the high vs low polymorphism tests.

#t.test comparisons of PWD_SC vs (KAZ, SKIVE, )
#MSM_SC vs MOLF_SC

@@

rand_SC.Length_MM.male.poly <- exactRLRT(lmer_MM_SC_length_male.poly)#this won't work either...hmm
```

@@

```{r long.sc_log.res_meanSC, eval=FALSE}
#hold off on reporting/ working with the long SC biv resutls

@@

#divide musc males into groups
long.SC.mouse.avs_4MM_male$Rec.group <- ifelse(grepl("PWD male", long.SC.mouse.avs_4MM_male$category), 1, 
                                    ifelse(grepl("MSM male", long.SC.mouse.avs_4MM_male$category), 1, 0))

@@

ToDo: make a melt.DF with the IFD1, IFD2, IFD3 in a single cols

@@

```{r IFD.old.plots, echo=FALSE, eval=FALSE}
#long SC Long_biv_mouse.avs_4MM
#real.long.bivData
#Only use 2CO bivalents
real.long.bivData.22 <- real.long.bivData[real.long.bivData$hand.foci.count == 2,]
real.long.bivData.22 <- real.long.bivData.22[(!is.na(real.long.bivData.22$hand.foci.count)),]

@@

The polymorphism frame-work

@@

logistic regression
$$  Rec \ Group ~=~ mouse \ average \ IFD $$
For these tests I am looking at the single IFD measure from 2CO bivalents.

```{r IFD.log.reg, echo=FALSE}
#
log.reg_IFD_ABS_mouse.av <- glm(Rec.group ~ mean_IFD.2CO_ABS, 
              data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),], family=binomial(link="logit"))

summary(log.reg_IFD_ABS_mouse.av)# 0.0366 *

log.reg_IFD_PER_mouse.av <- glm(Rec.group ~ mean_IFD.2CO_PER, 
                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),], family=binomial(link="logit"))

summary(log.reg_IFD_PER_mouse.av)#NS

### Single Bivalent levels

log.reg_IFD_ABS_single.biv <- glm(Rec.group ~ IFD1_ABS, 
              data=Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),], family=binomial(link="logit"))
summary(log.reg_IFD_ABS_single.biv)

log.reg_IFD_PER_single.biv <- glm(Rec.group ~ IFD1_PER, 
              data=Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),], family=binomial(link="logit"))
summary(log.reg_IFD_PER_single.biv)
```

```{r t.test_IFD_mouse.av, echo=FALSE, eval=FALSE}

#mouse.av - the mouse average IFD is from just 2CO
t.test_mouse.av.IFD_ABS <- t.test(
  mouse.avs_4MM$mean_IFD.2CO_ABS[(mouse.avs_4MM$Rec.group == 1)],

  mouse.avs_4MM$mean_IFD.2CO_ABS[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
  )

t.test_mouse.av.IFD_PER <- t.test(

  mouse.avs_4MM$mean_IFD.2CO_PER[(mouse.avs_4MM$Rec.group == 1)],

  mouse.avs_4MM$mean_IFD.2CO_PER[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
  )

```{r t.test_IFD,echo=FALSE}

#Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
#                                    ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))

#Curated_BivData <- add_IFD(Curated_BivData)
#THIS SHOULD ALREADY BE ADDED

#single biv level

Rec.group_IFD.ABS_test <- t.test(Curated_BivData$IFD1_ABS[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc") & (Curated_BivData$Rec.group ==0 ) & (Curated_BivData$hand.foci.count == 2 )],

     Curated_BivData$IFD1_ABS[ ( Curated_BivData$Rec.group ==1 ) & (Curated_BivData$hand.foci.count == 2 )]
)
Rec.group_IFD.ABS_test$p.value

## PER

Rec.group_IFD.PER_test <- t.test( Curated_BivData$IFD1_PER[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc") & (Curated_BivData$Rec.group ==0 )& (Curated_BivData$hand.foci.count == 2 )],

     Curated_BivData$IFD1_PER[ ( Curated_BivData$Rec.group ==1 ) &  (Curated_BivData$hand.foci.count == 2 )]
)#p-value = 0.04231

@@

#Single Biv Leve
single.biv.ABS.t.test <- t.test( Curated_BivData$IFD1_ABS[Curated_BivData$Rec.group == 1],

         Curated_BivData$IFD1_ABS[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc") &(Curated_BivData$Rec.group == 0)] )

single.biv.IFD_PER.t.test <- t.test( Curated_BivData$IFD1_PER[Curated_BivData$Rec.group == 1],

         Curated_BivData$IFD1_PER[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc") &(Curated_BivData$Rec.group == 0)] )
```

```{r show.IFD, echo=FALSE}

biv.IFDs_by_group <- ggplot(Curated_BivData[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc"),], aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS")+
  scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,100))
biv.IFDs_by_group

biv.IFDs_PER_by.group <- ggplot(Curated_BivData[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc"),], aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS \n Low=0, High=1")+
  scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
biv.IFDs_PER_by.group

oo.ifd <- ggplot(mouse.avs_4MM[ ( mouse.avs_4MM$sex == "male" ) & (mouse.avs_4MM$subsp == "Musc"),], aes(y = mean_IFD.2CO_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS")+
  scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = "free")+ylim(c(0,1))+theme(legend.position="none")

@@

Re-order the above figure if possible.

@@

```{r log.reg.CO1_pos, echo=FALSE}
#mouse av, norm.first foci is from 1CO bivalents
log.reg_1CO_pos_mouse.av <- glm(Rec.group ~ norm.first.foci.pos, 
                            data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
summary(log.reg_1CO_pos_mouse.av)

#single biv level
log.reg_1CO_pos_biv <- glm(Rec.group ~ PER_Foci_1, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
summary(log.reg_1CO_pos_biv)

```{r t.test_1CO_pos, echo=FALSE}

#mouse av
t.test_mouse.av.1CO_pos <- t.test(

  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$Rec.group == 1)],

  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
  )

#Single Biv Leve;
single.biv.1CO_pos_t.test <- t.test(

  Curated_BivData$PER_Foci_1[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1) & (Curated_BivData$Rec.group == 1)],

   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1) & (Curated_BivData$Rec.group == 0)]
  )

#0single biv level
```

the mouse average foci1 pos is more sig in t.test, but not log regression... (is something wrong?) Check the mouse averages for the F1_pos, there might be an outlier or mouse with v.few observations.

@@

```{r log.reg_siscoten, echo=FALSE}
#SISCOTEN
#mouse av, norm.first foci is from 1CO bivalents
log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,

                                 data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
summary(log.reg_SIS.co.ten_mouse.av)

#single biv level
log.reg_Sis.co.ten_biv <- glm(Rec.group ~ SisCoTen, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))

summary(log.reg_Sis.co.ten_biv)

```{r t.test_sis.co.ten,echo=FALSE}
#mouse av
t.test_mouse.av_siscoten <- t.test(

  mouse.avs_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],

  mouse.avs_4MM$mean.siscoten[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
  )

#Single Biv Level
t.test_single.biv_siscoten <- t.test(

  Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],

   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
  )
```

All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion struction/outcome.

@@

Mixed model result summary:

```{r log.reg_telo.dist, echo=FALSE}
#TELO.DIST
#mouse av, norm.first foci is from 1CO bivalents
log.reg_Telo.Dist_mouse.av <- glm(Rec.group ~ mean.telo.dist,

                                 data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
summary(log.reg_Telo.Dist_mouse.av)

#single biv level
log.reg_Telo.Dist_biv <- glm(Rec.group ~ telo_dist, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))

summary(log.reg_Telo.Dist_biv)

#PER versions
log.reg_Telo.Dist.PER_biv <- glm(Rec.group ~ telo_dist_PER, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))

summary(log.reg_Telo.Dist.PER_biv)
```

```{r t.test_Telo.Dist, echo=FALSE}

#mouse av
t.test_mouse.av_telo.dist <- t.test(

  mouse.avs_4MM$mean.telo.dist[(mouse.avs_4MM$Rec.group == 1)],

  mouse.avs_4MM$mean.telo.dist[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
  )

#Single Biv Level
single.biv_t.test_Telo.Dist <- t.test(

  Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],

   Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
  )

#Telo Dist PER
single.biv_t.test_Telo.Dist_PER <- t.test(

  Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],

   Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
  )

```{r cent.dist_show, echo=FALSE}
cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))

@@

```{r log.reg_CENT.dist, echo=FALSE}

#CENT.DIST
#mouse av, norm.first foci is from 1CO bivalents
log.reg_CENT.Dist_mouse.av <- glm(Rec.group ~ mean.cent.dist,

                                 data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
summary(log.reg_CENT.Dist_mouse.av)

#single biv level
log.reg_CENT.Dist_biv <- glm(Rec.group ~ dis.cent, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))

summary(log.reg_CENT.Dist_biv)

#PER versions
log.reg_CENT.Dist.PER_biv <- glm(Rec.group ~ dis.cent.PER, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))

summary(log.reg_CENT.Dist.PER_biv)
```

```{r t.test_CENT.Dist, echo=FALSE}
#CENT.DIST
#mouse av
t.test_mouse.av_Cent.dist <- t.test(

  mouse.avs_4MM$mean.cent.dist[(mouse.avs_4MM$Rec.group == 1)],

  mouse.avs_4MM$mean.cent.dist[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
  )

#Single Biv Level
single.biv_t.test_Cent.Dist <- t.test(

  Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],

   Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
  )

#Cent Dist PER
single.biv_t.test_Cent.Dist_PER <- t.test(

  Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],

   Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
  )
